---
title: "Differentially abundant species"
author: "Silvia Raineri"
date: "09/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.extra = '')
knitr::opts_knit$set(root.dir = "/Volumes/Mellor/Silvia/Microbiome/RATFAT04/") # NB substitute with appropriate working directory path
```

## Analysis of Differentially Abundant species 

This is an analysis of differentially abundant species in obese rats upon treatment long-term (42 Days) treatment with Sibutramine. To achieve this, DESeq2 was used with the model: ~ number + day

```{r load.libraries, echo=FALSE, message=FALSE}
source("config.R")
library('reshape2')
library("DESeq2")
library(tidyverse)
source('./R/microbiome_functions.R')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load species table 
species <- read.delim(species) 
species <- species[-1,] # take out unassigned
tax <- species$species

# filter to just keep the relevant samples (group H)
to.keep <- paste0("day3", "_", c("4", "14", "25", "33", "40", "55", "62", "75", "85"))
to.keep <- append(to.keep, paste0(gsub("day3", "day42", to.keep), "H"))

species <- species[,to.keep]
rownames(species) <- tax

# filter out low abundance species
rel.ab <- apply(species, 2, function(x) x / sum(x) )
f.relab <- rel.ab[rowSums(rel.ab >= 0.00001) >= 9,]
species <- species[rownames(species) %in% rownames(f.relab),]

# make metadata and run DESeq2
metadata <- data.frame(day=gsub("*_.*", "", to.keep),number= c("4", "14", "25", "33", "40",  "55", "62", "75", "85"))
rownames(metadata) <- to.keep
dds <- DESeqDataSetFromMatrix(species, colData=metadata, design= ~ number + day)
dds <- DESeq(dds, test='Wald', fitType='parametric')
res <- results(dds)

sig <- res[res$padj < 0.05 & !(is.na(res$padj)),]

cat("Number significant = ", nrow(sig))
rlog.dds <- rlog(dds)
```

# Plot Boxplots for all differentially abundant species 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
de.species.names <- rownames(sig[which(sig$padj <= 0.05),])
rlog <- assay(rlog.dds)
de.rlog <- rlog[match(de.species.names,rownames(rlog)),]

df <- data.frame(de.rlog, otus = rownames(de.rlog)) %>% melt()
df$variable <- gsub('X', '', df$variable)
df$Day <- gsub('_.*', '', df$variable)
df$number <- gsub('.*_', '', df$variable)
df$number <- gsub('H', '', df$number)
df$species <- gsub('.*s__', '', df$otus)

to.plot <- list(c(1:20), c(21:40),c(41:60), c(61:80), c(81:100), c(101:120), c(121:140), c(141:160), c(161:180), c(181:195))

for (i in 1:9){
  rows <- to.plot[[i]]
  print(rows)
  show <- rownames(de.rlog[rows,])
  dfm <- df[which(df$otus %in% show),]
  p <- ggboxplot(dfm, x = 'Day', y = 'value', color = 'species', group = "number") +
    geom_line(aes(group = dfm$number), color = "#ddd1db") +
    geom_point(aes(color = species), size = 2) +
    facet_wrap(~ species, scale="free_y") +
    labs(x = "Species", y = "Abundance") +
    theme(strip.text.x = element_text(size =6), axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size = 14), legend.position = "none")
  print(p)
}

```
# Focus on Bacteroides species 
Below a focus on all the differentially abundant Bacteroides species

```{r, echo=FALSE, message=FALSE, warning=FALSE}
show <- df$otus[grep("s__Bacteroides", df$otus)]
df2 <- df[df$otus %in% show,]

ggboxplot(df2, x = 'Day', y = 'value', color = 'species', group = "number") +
  geom_line(aes(group = df2$number), color = "#ddd1db") +
  geom_point(aes(color = species), size = 2) + 
  facet_wrap(~ species, scale="free_y") +
  labs(x = "Species", y = "Abundance") +
  theme(strip.text.x = element_text(size =8), axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size = 14), legend.position = "none")
```

# Ratio Bacteroidetes/Firmicutes 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
map <- readxl::read_xlsx('./Data/july19_map.xlsx')

# Extract total number of species for each phylum
Bact <- f.relab[grep("p__Bacteroidetes", rownames(f.relab)),]
Firmi <- f.relab[grep("p__Firmicutes", rownames(f.relab)),]

# calculate ratio
ratio <- colSums(Bact)/colSums(Firmi)
ratio.df <- data.frame(Samples = names(ratio), Ratio_Bacteroidetes_to_Firmicutes = ratio)
ratio.df$day <- gsub("_.*", "", ratio.df$Samples)
ratio.df$number <- gsub(".*_", "", ratio.df$Samples)
ratio.df$number <- gsub("H|A", "", ratio.df$number)
ratio.df$cohort <- if_else(ratio.df$number < 45, "1", "2")
ratio.df$group <- if_else(ratio.df$number %in% map$number[which(map$Group=="A")], "A", "H")
ratio.df$cohort[which(ratio.df$number ==8)] <- "1"

# Plot by cohort, and both cohorts together
bfc1 <- subset(ratio.df, cohort==1)
bfs1 <- ggboxplot(bfc1, x = 'day', y = 'Ratio_Bacteroidetes_to_Firmicutes', color = 'group') +
  stat_compare_means(method = 't.test', ref.group = 'day3', label = 'p.format', paired = T) +
  scale_color_manual(values= my_colors$Group[c('A', 'H')],labels = c("Ctrl", "Sib") ) +
  geom_line(aes(group = bfc1$number), color = "#ddd1db") +
  geom_point(aes(color = bfc1$group)) +
  facet_wrap(~ group)+
  scale_y_log10() +
  labs(title = 'Cohort 1', y = 'Ratio') +
  scale_x_discrete(labels = c('Day -3', 'Day 42'))  + theme(axis.text.x = element_text(angle = 90, hjust = 1),text = element_text(size=14))

bfc2 <- subset(ratio.df, cohort==2)
bfs2 <- ggboxplot(bfc2, x = 'day', y = 'Ratio_Bacteroidetes_to_Firmicutes', color = 'group') +
  stat_compare_means(method = 't.test', ref.group = 'day3', label = 'p.format', paired = T) +
  scale_color_manual(values= my_colors$Group[c('A', 'H')],labels = c("Ctrl", "Sib") ) +
  geom_line(aes(group = bfc2$number), color = "#ddd1db") +
  geom_point(aes(color = bfc2$group)) +
  facet_wrap(~ group)+
  scale_y_log10() +
  labs(title = 'Cohort 2', y = 'Ratio') +
  scale_x_discrete(labels = c('Day -3', 'Day 42'))  + theme(axis.text.x = element_text(angle = 90, hjust = 1),text = element_text(size=14))

bfs <- ggboxplot(ratio.df, x = 'day', y = 'Ratio_Bacteroidetes_to_Firmicutes', color = 'group') +
  stat_compare_means(method = 't.test', ref.group = 'day3', label = 'p.format', paired = T) +
  scale_color_manual(values= my_colors$Group[c('A', 'H')],labels = c("Ctrl", "Sib") ) +
  geom_line(aes(group = ratio.df$number), color = "#ddd1db") +
  geom_point(aes(color = ratio.df$group)) +
  facet_wrap(~ group)+
  scale_y_log10() +
  labs(title = 'Both cohorts', y = 'Ratio') +
  scale_x_discrete(labels = c('Day -3', 'Day 42'))  + theme(axis.text.x = element_text(angle = 90, hjust = 1),text = element_text(size=14))

l <- ggarrange(bfs1, bfs2, bfs, ncol= 3, nrow = 1, common.legend = T)
l
```

Below are the details of the R session that was used when this script was executed.

```{r sessionInfo}
print(sessionInfo())
```
