---
title: "Food Intake"
author: "Silvia Raineri"
date: "13/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.extra = '')
knitr::opts_knit$set(root.dir = "/Volumes/Mellor/Silvia/Microbiome/RATFAT04/") # NB substitute with appropriate working directory path
```

## Food Intake

Daily food (kJ) and water (g) intake were calculated for each animal. 
Food intake was calculated using the food energy values (20.79 kJ/g (high-fat chow), 23.44 kJ/g (chocolate) and 30.34 kJ/g (ground peanuts).

```{r load.libraries, echo=FALSE, message=FALSE}
source("config.R")
library('reshape2')
library(tidyverse)
source('./R/microbiome_functions.R')
```
# Combined food intake 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

map <- readxl::read_xlsx('./Data/july19_map.xlsx')

### Calculate Combined food intake 
download.file(food_intake_c1, "food_intake_c1.xlsx")
download.file(food_intake_c2, "food_intake_c2.xlsx")

hfd <- readxl::read_xlsx("food_intake_c1.xlsx", sheet = 1)
hfd2 <- readxl::read_xlsx("food_intake_c2.xlsx", sheet = 1)
hfd <- rbind(hfd, hfd2)

rats <- hfd$Rat
kj.hfd <- apply(hfd[,-1], 2, as.numeric)
kj.hfd <- apply(kj.hfd, 2, function(x) x  * 20.79)
rownames(kj.hfd) <- rats

choco <- readxl::read_xlsx("food_intake_c1.xlsx", sheet = 2)
choco2 <-  readxl::read_xlsx("food_intake_c2.xlsx", sheet = 2)
choco <- rbind(choco, choco2)

rats <- choco$Rat
kj.c <- apply(choco[,-1], 2, as.numeric)
kj.c <- apply(kj.c, 2, function(x) x  * 23.44)
rownames(kj.c) <- rats

peanuts <- readxl::read_xlsx("food_intake_c1.xlsx", sheet = 3)
peanuts2 <- readxl::read_xlsx("food_intake_c2.xlsx", sheet = 3)
peanuts <- rbind(peanuts, peanuts2)

rats <- peanuts$Rat
kj.p <- apply(peanuts[,-1], 2, as.numeric)
kj.p <- apply(kj.p, 2, function(x) x  * 30.34)
rownames(kj.p) <- rats

comb <- kj.hfd + kj.c + kj.p
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- melt(comb)
df$tr <- map$Group[match(df$Var1,map$Sample)]


dat <- list(
  FK506 = filter(df, tr %in% c('A', 'B')),
  Bupropion = filter(df, tr %in% c('A', 'C')),
  Naltrexone = filter(df, tr %in% c('A', 'D')),
  FK506_Bupropion = filter(df, tr %in% c('A', 'E')),
  FK506_Naltrexone  = filter(df, tr %in% c('A', 'F')),
  Bupropion_Naltrexone = filter(df, tr %in% c('A', 'G')),
  Sibutramine = filter(df, tr %in% c('A', 'H'))
) %>%
  bind_rows(.id = 'set')


ctrl <- filter(df, tr=="A")
ctrl <- ctrl %>% group_by(Var2) %>% summarize(value = mean(value, na.rm = T))
ctrl$tr <- "A"

d2 <- subset(dat, tr!="A")

ggplot() +
  geom_line(data=d2, aes(x = Var2, y = value, group = Var1, color = tr)) +
  geom_line(data=ctrl, aes(x = Var2, y = value, group = tr, color = tr), stat="summary", size = 1) +
  facet_wrap(~ set)+
  scale_color_manual(values = my_colors$Group) +
  scale_x_discrete(breaks = c("Day -3", "Day 0", "Day 15", "Day 36..43", "Day 42")) +
  labs(color = "Treatment", y = "kJ") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

# Stats
Statistical significance was assessed by ANOVA with the model: value ~ Day + Treatment + Day*Treatment. P-values were corrected using the Benjamini-Hochberg method.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("rstatix")
colnames(dat)<- c("set","sample", "Day", "value", "tr")

drugs <- c("FK506", "Bupropion", "Naltrexone", "FK506_Bupropion", "FK506_Naltrexone", "Bupropion_Naltrexone", "Sibutramine")

my.data <- list()
for(i in 1:7) {
  a <- subset(dat, set==drugs[i])
  my.data[[i]] <- a
}


# ANOVA results using Anova function from car package

anova.res.car <- list()
tukey.res.car <- list()
for (i in 1:7) {
  mod <- lm(value ~ Day + tr + Day*tr, my.data[[i]])
  x <- Anova(mod, type="II") 
  xt <- tukey_hsd(mod)
  anova.res.car[[i]] <- list(x)
  tukey.res.car[[i]] <- list(xt)
}

names(anova.res.car) <- drugs
print(anova.res.car)

p <- unlist(lapply(anova.res.car, function(x) x$p))
p.adj <- p.adjust(p, method = "BH")
formatC(p.adj,format="e")
```

# Tukey's post-hoc test on Sibutramine-treated samples (term= Day)
```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- tukey.res.car[[7]][[1]]
x <- subset(x, term=="Day") 
for (i in 1:length(x$group1)) {
  x$comparison[i] <- paste(x$group1[i],x$group2[i], collapse= "_")}

x$group1[which(x$group1=="Day 36...43")] <- "Day 36"
x$group1[which(x$group1=="Day 36...44")] <- "Day 37"
x$group2[which(x$group2=="Day 36...43")] <- "Day 36"
x$group2[which(x$group2=="Day 36...44")] <- "Day 37"


x$num <- as.numeric(gsub("Day", "", x$group1))
x$num2 <- as.numeric(gsub("Day", "", x$group2))

ggplot(x, aes(x = num, y =num2, color = padj)) +
  geom_point(color=if_else(x$p.adj<0.05, "red", "black"))+
  #facet_wrap(~ fast, scales = "free_x") +
  theme_classic() +
  theme(text = element_text(size=14),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #scale_x_discrete(breaks = x$num[which(x$p.adj<=0.05)]) +
  #scale_y_discrete(breaks = x$num[which(x$p.adj<=0.05)]) +
  labs(x = "First term (Day) of Comparison", y = "Second term (Day) of comparison")
```

# Food intake composition

```{r, echo=FALSE, message=FALSE, warning=FALSE}
h <- data.frame(rats, kj.hfd, diet = paste("hfd"))
ch <- data.frame(rats, kj.c,  diet = paste("chocolate"))
p <- data.frame(rats, kj.p, diet = paste("peanuts"))

diet <- rbind(h, ch, p)  %>% melt(id.vars=c("rats", "diet"))
diet$tr <- map$Group[match(diet$rats,map$Sample)]

d2 <- diet %>% group_by( rats, variable) %>% 
  mutate(percent = value / sum(value) * 100)

ggplot(d2, aes(x = variable, y = percent, fill = diet)) +
  geom_bar(stat= "summary", position = 'fill') + 
  scale_fill_manual(values = c('#1e87d2','#542a0c', '#db874a')) +
  facet_wrap(~ tr) +
  labs(title = 'Diet composition by Treatment Group (%)', y = "% of Diet")+
  scale_x_discrete(breaks = c("Day..3", "Day.0", "Day.15", "Day.36..43", "Day.42")) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_bw() +
  theme(text = element_text(size=14),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
Below are the details of the R session that was used when this script was executed.

```{r sessionInfo}
print(sessionInfo())
```