---
title: "Weight Data"
author: "Silvia Raineri"
date: "1 March 2018"
output: html_document
---
## Weight Data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.extra = '')
knitr::opts_knit$set(root.dir = "/Volumes/Mellor/Silvia/Microbiome/RATFAT04/") # NB substitute with appropriate working directory path
```

As the hypothesis is that the drug is acting on these rats' weight, let's see how this parameter has changed throughout the experiment. To do so, we will consider both cohorts. 

```{r load.libraries, echo=FALSE, message=FALSE}
source("config.R")
library('reshape2')
library("DESeq2")
library(tidyverse)
source('./R/microbiome_functions.R')
``` 

```{r, echo = F, error = F, message = F, warning = F}
# All weights
weight.chart <- read.delim(weight, sep = ',')

#Create a custom color scale
library(RColorBrewer)
myColors <- c('#a8c0fc','#ff9b9b', '#fccd76', '#55d2cc', '#d43a5b', '#ffa227','#359c7a', '#1b2152')
names(myColors) <- levels(weight.chart$Group)
colScale <- scale_colour_manual(name = "grp",values = myColors)

weight.chart$Treatment <- factor(weight.chart$Treatment, levels = c('Ctrl', 'FK506', 'Bupropion', 'Naltrexone','FK506_B', 'FK506_N', 'B_N', 'Sibutramine'))
weight.chart <- arrange(weight.chart, Treatment)

```


```{r, echo = F, error = F, message = F, warning = F}
ggplot(data = weight.chart, aes(x = Timepoint, y = Weight.g., group = Sample,  col = Group)) +
  geom_line(aes(linetype = Cohort)) +
  geom_text(aes(label=weight.chart$Sample),hjust=1.2, vjust=0.01, size = 1) +
  scale_colour_manual(name = "Group",values = myColors) +
  facet_wrap(~ Treatment, nrow = 2) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) 

```
  

ANOVA was used to calculate any significant differences between the weight of the rats with the model: weight ~ timepoint + Cohort.  

```{r, echo = FALSE,fig.height= 7, fig.width=7,error = F, message = F, warning = F}

weight <- readxl::read_xlsx('./Data/RATFAT04_weights.xlsx') %>% gather( 'W_-3':'W_42', key = 'timepoint', value = 'weight' )
weight$weight <- gsub(",",".",weight$weight) %>% as.numeric(.)
sub <- subset(weight, timepoint=="W_-3" | timepoint=="W_42")

### ANOVA ---------------------
groups <- c("A", "B", "C", "D", "E", "F", "G", "H")
my.data <- list()
for(i in 1:8) {
  a <- subset(sub, Group==groups[i])
  my.data[[i]] <- a
}

anova.res.car <- list()
for (i in 1:8) {
  mod <- lm(weight ~ timepoint + Cohort, my.data[[i]])
  x <- car::Anova(mod, type="II") 
  xt <- rstatix::tukey_hsd(mod)
  anova.res.car[[i]] <- list(x, xt)
}

names(anova.res.car) <- groups
print(anova.res.car)
```
# Alpha diversity 

```{r, echo = FALSE,fig.height= 7, fig.width=7, error = F, message = F, warning = F}
library("phyloseq")
source('./R/microbiome_functions.R')
map <- readxl::read_xlsx('./Data/july19_map.xlsx')

species <- read.delim(species) # NB! No species data about samples   '48'  '50'  '54'  '50H' 
colnames(species) <- gsub('day42_', '',colnames(species))
colnames(species) <- gsub('day3_', '', colnames(species))
species_id <- species[,1]

map <- map[which(map$Sample %in% colnames(species)),]
map <- as.data.frame(map)
rownames(map) <- map$Sample

species <- species[, match(map$Sample, colnames(species))]
rownames(species) <- species_id


tax <- read.table(tax, header = T)
tax <- as.matrix(tax)
rownames(tax) <- species_id


physeq <- make_phyloseq(species[-1,], tax[-1,], map)
```
```{r, echo = FALSE,fig.height= 5, fig.width=7, error = F, message = F, warning = F}
otu_table(physeq) <- otu_table(round(as((otu_table(physeq)), 'matrix')), taxa_are_rows(physeq))
alpha.data <- estimate_richness(otu_table(physeq), measures = c('Simpson', 'Chao1'))
rownames(alpha.data) <- map$Sample
alpha.df <- data.frame(alpha.data, weight = map$weight_loss[match(rownames(alpha.data), map$Sample)], Group = map$Group[match(rownames(alpha.data), map$Sample)], Day = map$Day[match(rownames(alpha.data), map$Sample)],
                       Weight = map$Weight[match(rownames(alpha.data), map$Sample)], cohort = map$cohort[match(rownames(alpha.data), map$Sample)], number = map$number[match(rownames(alpha.data), map$Sample)])
alpha.df$Day <- factor(alpha.df$Day, levels =c('Pre', 'Day_42'))
alpha.df <- alpha.df[-which(alpha.df$number == 54 | alpha.df$number == 48),]

my_labs <- c('Ctrl', 'FK506', 'Bupropion', 'Naltrexone', 'FK506_B', 'FK506_N', 'Bupropion_Naltrexone', 'Sibutramine')

ggboxplot(alpha.df, x = 'Day' , y = 'Chao1', color = 'Group') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size=14)) +
  geom_line(aes(group = number), color = '#ddd1db') +
  facet_wrap(~ Group, scales = 'free_y', nrow =2)+
  geom_point(aes(color = Group)) +
  #coord_cartesian(ylim = c(NA,1.05)) +
  scale_color_manual(values = my_colors$Group, labels = my_labs ) 

dat_paired <- list(
  FK506 = filter(alpha.df, Group %in% c('A', 'B')),
  Bupropion = filter(alpha.df, Group %in% c('A', 'C')),
  Naltrexone = filter(alpha.df, Group %in% c('A', 'D')),
  FK506_Bupropion = filter(alpha.df, Group %in% c('A', 'E')),
  FK506_Naltrexone  = filter(alpha.df, Group %in% c('A', 'F')),
  Bupropion_Naltrexone = filter(alpha.df, Group %in% c('A', 'G')),
  Sibutramine = filter(alpha.df, Group %in% c('A', 'H'))
) %>%
  bind_rows(.id = 'set')
dat_paired$set <- factor(dat_paired$set, levels = c('FK506','Bupropion', 'Naltrexone', 'FK506_Bupropion', 'FK506_Naltrexone', 'Bupropion_Naltrexone', 'Sibutramine'))
names(my_colors$Group)

ggplot(dat_paired, aes(x = Group, y = Chao1)) +
  #stat_compare_means(method = 't.test', label = 'p.format', ref.group = 'A') + 
  geom_boxplot(aes(color = Group)) +
  scale_color_manual(values = my_colors$Group, labels = my_labs)+
  facet_wrap(~ set + Day, scales = 'free') +
  geom_point(aes(color = Group)) +
  #geom_text(data = ptest, aes(x = 1, y = 0.4, label = paste0("p=",round(p.value,3)))) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size=13), legend.position = "bottom") +
  guides(color = guide_legend(ncol = 2)) 

## Adjusted P-values alpha diversity
ptest <- dat_paired %>% group_by(set, Day) %>% summarize(p.value = t.test(Simpson ~ Group)$p.value)
simpson.group <- data.frame(ptest, adj = p.adjust(ptest$p.value, method = "BH"))

ptest <- dat_paired %>% group_by(set, Day) %>% summarize(p.value = t.test(Chao1 ~ Group)$p.value)
chao1.group <- data.frame(ptest, adj = p.adjust(ptest$p.value, method = "BH"))

ptest <- alpha.df %>% group_by(Group) %>% summarize(p.value = t.test(Chao1 ~ Day, paired = TRUE, alternative = "two.sided")$p.value)
chao1.time <- data.frame(ptest, adj = p.adjust(ptest$p.value, method = "BH"))

ptest <- alpha.df %>% group_by(Group) %>% summarize(p.value = t.test(Simpson ~ Day, paired = TRUE, alternative = "two.sided")$p.value)
simpson.time <- data.frame(ptest, adj = p.adjust(ptest$p.value, method = "BH"))

```

# Beta Diversity 

```{r, echo = FALSE,error = F, message = F, warning = F}

physeq.relab <- transform_sample_counts(physeq, function(x) x / sum(x) )
physeq.rel.filtered <- filter_taxa(physeq.relab, function(x) sum(x >= 0.001) >= 10, TRUE)
physeq.filtered <- subset(otu_table(physeq), rownames(otu_table(physeq)) %in% rownames(otu_table(physeq.rel.filtered)))
physeq.filtered <- merge_phyloseq(physeq.filtered, tax_table(physeq), sample_data(physeq))

################################
### Function: get.betadiv.df ###
################################
get.betadiv.df <- function(physeq, Group) {
  # obtain data frame to plot beta diversity PCA plots of each treatment vs ctrl
  # 1. Subset physeq  and map into Ctrl and group of interest 
  sub <- data.frame(sample_data(physeq)[which(sample_data(physeq)$Group=="A" |sample_data(physeq)$Group==Group),])
  otu <- otu_table(physeq)[, match(sub$Sample,colnames(otu_table(physeq)))]
  print(dim(sample_data(sub)))
  print(dim(otu_table(otu)))
  # 2. calculate beta diversity
  bdiv <- transform_sample_counts(otu, function(x) 1E6 * x/sum(x))
  print(class(bdiv))
  dist.matrix <- t(data.frame(otu_table(bdiv)))
  bdiv <- vegdist(dist.matrix, method = "bray")
  pc <- cmdscale(bdiv,k=2)
  rownames(pc)<-gsub("X", "", rownames(pc))
  my.df <- data.frame(V1 =pc[,1], V2= pc[,2],  
                      ID = map$Sample[match(rownames(pc), map$Sample)], 
                      treatment = map$Treatment[match(rownames(pc), map$Sample)], 
                      timepoint = map$Day[match(rownames(pc), map$Sample)], 
                      group = map$Group[match(rownames(pc), map$Sample)])
  # 3. Adonis
  print(adonis(bdiv ~ Treatment * Group + cohort, data = data.frame(sub)))
  return(my.df)
}


fk506 <- get.betadiv.df(physeq.filtered,'B')
bupropion <- get.betadiv.df(physeq,'C')
naltrexone <- get.betadiv.df(physeq,'D')
fk506_bupropion <- get.betadiv.df(physeq,'E')
fk506_naltrexone <- get.betadiv.df(physeq,'F')
bupropion_naltrexone <- get.betadiv.df(physeq,'G')
sibutramine <- get.betadiv.df(physeq,'H')
  
my.plot <- bind_rows(fk506, bupropion, naltrexone, fk506_bupropion, fk506_naltrexone, bupropion_naltrexone, sibutramine, .id = 'set')
set.labs <- c('FK506','Bupropion', 'Naltrexone', 'FK506_Bupropion', 'FK506_Naltrexone', 'Bupropion_Naltrexone', 'Sibutramine')
names(set.labs) <- c("1", "2", "3", "4", "5", "6", "7")

ggplot(my.plot, aes(x = V1, y = V2, color = group)) +
  geom_point(stat='identity', size = 2.5, aes(shape = timepoint)) +
  scale_color_manual(values = my_colors$Group) +
  facet_wrap(~ set , labeller = labeller(set = set.labs)) +
  labs(x= 'PC1', y = 'PC2',  title = 'Beta Diversity', color = "Treatment", shape = "Timepoint") +
  theme_bw()

dat_text <- data.frame(
  label = c("R2 = 0.02 (ns)", "R2 = 0.06 (ns)", "R2 = 0.033 (ns)", "R2 = 0.05 (ns)", "R2 = 0.038 (ns)", "R2 = 0.13 (***)", "R2 = 0.23 (***)"),
  set   = c("1", "2", "3", "4", "5", "6", "7")
)


```



Below are the details of the R session that was used when this script was executed.

```{r sessionInfo}
print(sessionInfo())
```

